Aspects of RDF Data Constraints in the Social, Behavioural and Economic Sciences

### 2015-04-30 paper/paper7/paper7.tex

For research institutes, data libraries, and data archives,
RDF data validation according to predefined constraints is a much sought-after feature, 
particularly as this is taken for granted in the XML world.
Based on our work in the DCMI RDF Application Profiles Task Group and in cooperation with the W3C Data Shapes Working Group, we identified and published by today 82 types of constraints that are required by various stakeholders for data applications.
In this paper, we formulate 246 constraints of 53 different types on six different vocabularies (Disco, QB, SKOS, PHDD, DCAT, XKOS) and classify them according to the complexity level of their type and their severity level. For 114 of these constraints, we evaluate the data quality of 15,694 data sets (4.26 billion triples) of research data for the social, behavioural, and economic (SBE) sciences obtained from 33 SPARQL endpoints. 
Based on the results, we formulate several hypotheses to direct the further development of constraint languages.

@@

The modeling languages \emph{RDF}, \emph{RDFS}, and \emph{OWL} are typically used to define vocabularies.
\emph{Vocabulary constraint types}
denotes the set of constraint types whose constraints 
can be extracted completely automatically out of the formal specifications of vocabularies.
As vocabularies have been specified using \emph{RDF}, \emph{RDFS}, and \emph{OWL},
\emph{vocabulary constraints} ensure that the data is consistent with the intended syntaxes, semantics, and integrity of vocabularies' data models.
%The {\em existential quantification} (\emph{R-86}) 
%\ms{Study $\equiv$ $\exists$ product.LogicalDataSet} (\emph{DL}), e.g., restricts studies to contain at least one data set which is expressible in OWL 2:
\emph{Minimum qualified cardinality restrictions} (\emph{R-74}), e.g., guarantee that individuals of given classes are connected by particular properties to at least n different individuals/literals of certain classes or data ranges.
This way, it is expressible in \emph{OWL 2} that a \emph{phdd:TableStructure} has (\emph{phdd:column}) at least one \emph{phdd:Column}:
\ke{Ich dachte, das kommt aus dem Vokabular. Dann muss das heißen: For example, in PHDD, a minimum qualified cardinality constraint can be obtained from the OWL definition of this restriction class, ensuring that a \emph{phdd:TableStructure} has (\emph{phdd:column}) at least one \emph{phdd:Column}:}
\begin{ex}
[   a owl:Restriction ; rdfs:subClassOf TableStructure ;
    owl:minQualifiedCardinality 1 ;
    owl:onProperty column ;
    owl:onClass Column ] .
\end{ex}

\emph{Simple} and \emph{complex constraints} are in contrast to \emph{vocabulary constraints} not explicitly defined within formal specifications of vocabularies.
\emph{Simple} and \emph{complex constraints} are defined according to textual descriptions of the intended semantics of vocabularies.  
\emph{Simple constraint types} is the set of constraint types whose constraints can be easily defined without much effort in addition to \ke{einfügen: the already defined (auf den Lesefluss achten, das ist eine sehr pragmatische klassifikation und muss nicht übertrieben formal und trocken geschrieben sein)} \emph{vocabulary constraints}.
{\emph{Data property facets} (\emph{R-46}) is an example of a \emph{simple constraint type} which enables to declare frequently needed facets for data properties in order to validate input against simple conditions including min/max values, regular expressions, and string length.
The abstract of series/studies, e.g., should have a minimum length.

\emph{Complex constraint types} encompass constraint types for which the definition of constraints is rather complex and cannot be derived from vocabulary definitions.
For assessing the quality of thesauri, e.g., we concentrate on the graph-based structure and apply graph- and network-analysis techniques.
An example of such constraints of the constraint type \emph{structure} is that 
a thesaurus should not contain many orphan concepts, i.e., concepts without any associative or hierarchical relations, lacking context information valuable for search. 
\emph{Complex constraints} show the importance to develop constraint languages enabling to describe more complex constraints.\ke{der satz macht so keinen sinn... aber er ist ein weiteres gutes beispiel, warum du viel mehr auf die formulierungen achten musst. Ich weiß genau was du meinst, du hast auch recht, aber hier steht trotzdem ein sinnloser satz, so wie er formuliert ist.}

\ke{ich würde da auch für simple und vor allem für complex konkrete beispiele in code einfügen}

@@

\ke{da muss mindestens ein überleitender satz rein, der noch mal den unterschied von constraint types und types aufgreift und erklärt, dass es jetzt um konkrete constraints auf den vokabularen geht, deren severity eben vom konkreten kontext abhängt und nicht allgemein über den type bestimmt werden kann.}

\emph{SBE} experts determined the default \textbf{\emph{severity level}} (\emph{R-158}) for each of the 246 constraints to indicate how serious the violation of the constraint is.
We propose an extensible metric to measure the continuum of severity levels ranging from \emph{informational} to \emph{error}.\ke{oder etwas weniger anmaßend: we use the commonly accepted classification of log messages in software development and distingush informational, warning and error.}
According to the default severity level of constraints, the complete set of constraints encompasses three disjoint \emph{sets of constraints}:\ke{auch dieser satz macht keinen sinn für mich, ebensowenig wie die extrem erhellende auflistung, die gleich folgt... mein vorschlag von eben kann das komplett ersetzen.}
\begin{itemize}
	\item \textbf{\emph{informational constraints}}: constraints with severity level \emph{informational}
	\item \textbf{\emph{warning constraints}}: constraints with severity level \emph{warning}
	\item \textbf{\emph{error constraints}}: constraints with severity level \emph{error}
\end{itemize}
Violations of \emph{informational constraints} point to desirable but not necessary data improvements to achieve RDF representations which are ideal in terms of syntax and semantics of used vocabularies. 
Data not conforming to \emph{warning} and \emph{error constraints} is syntactically and/or semantically not correctly represented.
Data not conforming to \emph{warning constraints} but to \emph{error constraints} could be processed further.
Data not corresponding to \emph{error constraints}, however, cannot be processed further after validation. 
As the purpose of \emph{vocabulary constraints} is to ensure explicitly stated semantics of vocabularies, 
their default severity levels are in most cases very strong (\emph{error}) and 
in average stronger than the severity levels of \emph{simple} and \emph{complex constraints}.
As a consequence, violating many \emph{vocabulary constraints} is an indicator for bad (meta)data quality\footnote{For simplicity reasons, we only assign severity levels to \emph{vocabulary constraints} in this paper in case they differ from \emph{error constraints}.}.
\ke{den ganzen letzten absatz finde ich unübersichtlich. Kann man das nicht klarer sagen: Error sind syntaktische oder semantische fehler, die zu einem abbruch der datenverarbeitung führen sollten. warnings sind syntaktische oder semantische fehler, die aber normalerweise nicht zu einem abbruch führen müssen. verletzungen von informational constraints weisen auf potentielle verbesserungen in den daten hin, werden aber nicht als fehler angesehen (der satz war in ordnung). die fussnote ist doch vollkommen fehl am platz, oder? es werden ja nicht alle constraints aufgelistet und wie ihr das implementiert habt und ob ihr euch arbeit erspart habt weil ihr default severity bei bestimmten constraint types vorgebt, das interessiert doch keinen.}

@@

The constraint type \emph{vocabulary} guarantees that users do not invent new or use deprecated terms of vocabularies.
%Out-dated classes and properties of previous vocabulary versions can be marked as deprecated.
%The constraint types \emph{context-specific valid classes and properties} (\emph{R-209; R-210}) can be used to specify which classes and properties are valid in which context - here a given vocabulary version.
\emph{Value is valid for datatype} (\emph{R-223}) constraints serve to make sure that all literal values are valid with regard to their datatypes - as stated in the vocabularies.
Thus, it is checked that all date values (e.g., {{\em dcterms:date}) are actually of the datatype \emph{xsd:date} and that \emph{xsd:nonNegativeInteger} values (e.g. \emph{disco:frequency}) are not negative.
Depending on property datatypes, two different literal values have
a specific ordering with respect to operators like \textless  (\emph{R-43: literal value comparison}).
Start dates (\emph{disco:startDate}), e.g., must be before (\emph{\textless}) end dates (\emph{disco:endDate}).

@@

As vocabulary constraints are contained within formal specifications of vocabularies and therefore directly ensure their syntax and semantics,
violations of vocabulary constraints are more severe than violations caused by simple constraints which are not part of vocabularies.
In many cases, complex constraints are also needed to express severe constraints which cannot be expressed by modeling languages.
The relation between complexity and severity shows the importance to develop constraint languages which are enable to express severe complex constraints.

@@

\tb{ToDO}

>>>>

For research institutes, data libraries, and data archives,
RDF data validation according to predefined constraints is a much sought-after feature, 
particularly as this is taken for granted in the XML world.
Based on our work in the DCMI RDF Application Profiles Task Group and in cooperation with the W3C Data Shapes Working Group, we identified and published by today 82 types of constraints that are required by various stakeholders for data applications.
In this paper, we formulate 246 constraints of 53 different types on six different vocabularies (Disco, QB, SKOS, PHDD, DCAT, XKOS) and classify them according to the complexity level of their type and their severity level. For 114 of these constraints, we evaluate the data quality of 15,694 data sets (4.26 billion triples) of research data for the social, behavioural, and economic (SBE) sciences obtained from 33 SPARQL endpoints. \ke{check numbers}
Based on the results, we formulate several hypotheses to direct the further development of constraint langugaes.

@@

The modeling languages \emph{RDF}, \emph{RDFS}, and \emph{OWL} are typically used to formally specify vocabularies.
\emph{Vocabulary constraint types}
denotes the set of constraint types whose constraints 
can be extracted completely automatically out of formal specifications of vocabularies.
As vocabularies have been specified using \emph{RDF}, \emph{RDFS}, and \emph{OWL},
\emph{vocabulary constraints} ensure that the data is consistent with the intended syntaxes, semantics, and integrity of vocabularies' data models.
%The {\em existential quantification} (\emph{R-86}) 
%\ms{Study $\equiv$ $\exists$ product.LogicalDataSet} (\emph{DL}), e.g., restricts studies to contain at least one data set which is expressible in OWL 2:
\emph{Minimum qualified cardinality restrictions} (\emph{R-74}), e.g., guarantee that individuals of given classes are connected by particular properties to at least n different individuals/literals of certain classes or data ranges.
In \emph{PHDD}, a \emph{minimum qualified cardinality restriction} can be obtained from the OWL definition of this restriction class, ensuring that a \emph{phdd:TableStructure} has (\emph{phdd:column}) at least one \emph{phdd:Column}:
\begin{ex}
[   a owl:Restriction ; rdfs:subClassOf TableStructure ;
    owl:minQualifiedCardinality 1 ;
    owl:onProperty column ;
    owl:onClass Column ] .
\end{ex}

\emph{Simple} and \emph{complex constraints} are in contrast to \emph{vocabulary constraints} not explicitly defined within formal specifications of vocabularies.
\emph{Simple} and \emph{complex constraints} are defined according to textual descriptions of the intended semantics of vocabularies.  
\emph{Simple constraint types} is the set of constraint types whose constraints can be easily defined without much effort in addition to the already defined \emph{vocabulary constraints}.
{\emph{Data property facets} (\emph{R-46}) is an example of a \emph{simple constraint type} which enables to declare frequently needed facets for data properties in order to validate input against simple conditions including min/max values, regular expressions, and string length.
The abstract of series/studies, e.g., should have a minimum length which is easily and concisely expressible by SPARQL: 
\begin{ex}
SELECT ?study WHERE {
    ?study dcterms:abstract ?abstract . FILTER ( STRLEN ( ?abstract ) = 10 ) . }
\end{ex}

\emph{Complex constraint types} encompass constraint types for which the definition of constraints is rather complex and cannot be derived from vocabulary definitions.
For assessing the quality of thesauri, e.g., we concentrate on the graph-based structure and apply graph- and network-analysis techniques.
An example of such constraints of the constraint type \emph{structure} is that 
a thesaurus should not contain many orphan concepts, i.e., concepts without any associative or hierarchical relations, lacking context information valuable for search. This \emph{complex constraint} is only expressible by \emph{SPARQL} and not directly understandable:
\begin{ex}
SELECT ?concept WHERE {
    ?concept a [rdfs:subClassOf* skos:Concept] .
    FILTER NOT EXISTS { ?concept ?p ?o . 
        FILTER ( ?p IN ( skos:related, skos:relatedMatch, skos:broader, ... ) ) . }
\end{ex}
\emph{Complex constraints} are in most cases only expressible by plain \emph{SPARQL}
which shows the importance to develop suitable constraint languages.

@@

A concrete constraint is instantiated from one of the 82 constraint types and is defined for a specific vocabulary.
It does not make sense to determine the severity of constraint violations of an entire constraint type,
as the severity of the violation of a given constraint depends on the individual context and vocabulary.
%as the violation of one constraint may be severe,
%whereas, the violation of another constraint of the same type may be minor.
\emph{SBE} experts determined the default \textbf{\emph{severity level}} (\emph{R-158}) for each of the 246 constraints to indicate how serious the violation of the constraint is.
%We propose an extensible metric to measure the continuum of severity levels ranging from \emph{informational} to \emph{error}.
We use the commonly accepted classification of log messages in software development and distinguish \emph{informational}, \emph{warning} and \emph{error}.
%According to the default severity level of constraints, the complete set of constraints encompasses three disjoint \emph{sets of constraints}:\ke{auch dieser satz macht keinen sinn für mich, ebensowenig wie die extrem erhellende auflistung, die gleich folgt... mein vorschlag von eben kann das komplett ersetzen.}
%\begin{itemize}
	%\item \textbf{\emph{informational constraints}}: constraints with severity level \emph{informational}
	%\item \textbf{\emph{warning constraints}}: constraints with severity level \emph{warning}
	%\item \textbf{\emph{error constraints}}: constraints with severity level \emph{error}
%\end{itemize}
Violations of \emph{informational constraints} point to desirable but not necessary data improvements to achieve RDF representations which are ideal in terms of syntax and semantics of used vocabularies. 
They are not seen as serious errors.
\emph{Errors} are syntactic or semantic errors which should cause the abortion of data processing. 
\emph{Warnings} are syntactic or semantic errors which typically should not lead to an abortion of data processing.
%Data not conforming to \emph{warning} and \emph{error constraints} is syntactically and/or semantically not correctly represented.
%Data not conforming to \emph{warning constraints} but to \emph{error constraints} could be processed further.
%Data not corresponding to \emph{error constraints}, however, cannot be processed further after validation. 
As the purpose of \emph{vocabulary constraints} is to ensure explicitly stated semantics of vocabularies, 
their default severity levels are in most cases very strong (\emph{error}) and 
in average stronger than the severity levels of \emph{simple} and \emph{complex constraints}.
As a consequence, violating many \emph{vocabulary constraints} is an indicator for bad (meta)data quality.

@@

The constraint type \emph{vocabulary} guarantees that users do not invent new or use deprecated terms of vocabularies.
%Out-dated classes and properties of previous vocabulary versions can be marked as deprecated.
%The constraint types \emph{context-specific valid classes and properties} (\emph{R-209; R-210}) can be used to specify which classes and properties are valid in which context - here a given vocabulary version.
\emph{Value is valid for datatype} (\emph{R-223}) constraints serve to make sure that all literal values are valid with regard to their datatypes - as stated in the vocabularies.
Thus, it is checked that all date values (e.g., {{\em dcterms:date}) are actually of the datatype \emph{xsd:date} and that \emph{xsd:nonNegativeInteger} values (e.g. \emph{disco:frequency}) are not negative\footnote{For simplicity reasons, we only assign severity levels to \emph{vocabulary constraints} if they differ from \emph{error}.}.
Depending on property datatypes, two different literal values have
a specific ordering with respect to operators like \textless  (\emph{R-43: literal value comparison}).
Start dates (\emph{disco:startDate}), e.g., must be before (\emph{\textless}) end dates (\emph{disco:endDate}).

@@

As \emph{vocabulary constraints} are part of formal specifications of vocabularies,
violations of \emph{vocabulary constraints} are more severe than violations caused by \emph{simple constraints}.
As \emph{SPARQL} is still used to express \emph{complex constraints} and as 44\% of all \emph{complex constraints} are serious,
there is an obvious need to develop constraint languages which enable to express \emph{complex constraints}.

@@
